Sweat electrolytes and CFTR activity in the era of CFTR modulators

### 2025-08-07 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
genotype severity analysis

---
title: "Sweat electrolytes and CFTR activity in the era of CFTR modulators"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 10
    toc-float: true
    number-sections: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
library(knitr)
library(Hmisc)
library(lubridate)
library(readxl)
library(tidyverse)
library(arsenal)
library(pROC)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Variability",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

@@

@@

@@

@@

@@

## Histograms and Density Plots

```{r}
#| warning: false
ggplot(df, aes(x = Cl, fill = Diagnosis)) +
  geom_histogram(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
  geom_histogram(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
  geom_histogram(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
  theme_bw() +
  theme(axis.title.y = element_blank())
ggplot(df, aes(x = Cl, fill = Diagnosis)) +
  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
  theme_bw() +
  theme(axis.title.y = element_blank())
```

```{r}
#| warning: false
ggplot(df, aes(x = Na, fill = Diagnosis)) +
  geom_histogram(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
  geom_histogram(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
  geom_histogram(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
  theme_bw() +
  theme(axis.title.y = element_blank())
ggplot(df, aes(x = Na, fill = Diagnosis)) +
  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
  theme_bw() +
  theme(axis.title.y = element_blank())
```

```{r}
#| warning: false
ggplot(df, aes(x = `Na:Cl`, fill = Diagnosis)) +
  geom_histogram(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
  geom_histogram(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
  geom_histogram(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
  theme_bw() +
  theme(axis.title.y = element_blank())
ggplot(df, aes(x = `Na:Cl`, fill = Diagnosis)) +
  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
  theme_bw() +
  theme(axis.title.y = element_blank())
ggplot(df, aes(x = `Na:Cl`, fill = Diagnosis)) +
  geom_density(data = subset(df, Diagnosis == "Non-CF"), alpha = 0.2) +
  geom_density(data = subset(df, Diagnosis == "CFTR-RD or CRMS"), alpha = 0.2) +
  geom_density(data = subset(df, Diagnosis == "Cystic Fibrosis"), alpha = 0.2) +
  xlim(0, 2) +
  theme_bw() +
  theme(axis.title.y = element_blank())
```

>>>>

---
title: "Sweat electrolytes and CFTR activity in the era of CFTR modulators"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 10
    toc-float: true
    number-sections: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(knitr)
library(Hmisc)
library(lubridate)
library(readxl)
library(tidyverse)
library(arsenal)
library(pROC)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Variability",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
source(paste0(github_dir, "/CF/genotype_class_severity.R"))
knitr::opts_knit$set(root.dir = home_dir)
```

@@

# Get genotype severity
## By severity
df$geno_risk_1 <- cf_genotype_class_severity(df$Genotypes1)
df$geno_risk_2 <- cf_genotype_class_severity(df$Genotypes2)
df$geno_risk_1[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno_risk_2[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno_risk <- paste(df$geno_risk_1, df$geno_risk_2, sep = "/")
df$geno_risk <- factor(df$geno_risk,
  levels = c(
    "Mild/Mild", "Mild/Severe", "Mild/Unknown", "Severe/Mild", "Severe/Severe",
    "Severe/Unknown", "Unknown/Severe", "Unknown/Unknown", "Not CF/Not CF"
  ),
  labels = c("M/M", "M/S", "M/U", "M/S", "S/S", "S/U", "S/U", "U", "Not CF")
## By function
df$geno1_class <- cf_genotype_class_severity(df$Genotypes1, out = "class")
df$geno1_class[grep("508", df$Genotypes1)] <- "F508del"
df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("I", "II", "III"), "MF")
df$geno1_class <- replace(df$geno1_class, df$geno1_class %in% c("IV", "V"), "RF")
df$geno2_class <- cf_genotype_class_severity(df$Genotypes2, out = "class")
df$geno2_class[grep("508", df$Genotypes2)] <- "F508del"
df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("I", "II", "III"), "MF")
df$geno2_class <- replace(df$geno2_class, df$geno2_class %in% c("IV", "V"), "RF")
df$geno1_class[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno2_class[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno_class <- paste(df$geno1_class, df$geno2_class, sep = "/")
df$geno_class <- factor(df$geno_class,
  levels = c(
    "F508del/F508del", "F508del/MF", "F508del/RF", "F508del/Unknown",
    "MF/F508del", "MF/MF", "MF/RF", "MF/Unknown", "RF/F508del", "RF/MF",
    "RF/Unknown", "Unknown/F508del", "Unknown/MF", "Unknown/Unknown",
    "Not CF/Not CF"
  ),
  labels = c(
    "MF/MF", "MF/MF", "MF/RF", "MF/U", "MF/MF", "MF/MF", "MF/RF", "MF/U",
    "MF/RF", "MF/RF", "RF/U", "MF/U", "MF/U", "U", "Not CF"
# Organize columns
  select(Name:Genotypes2, geno_risk, geno_class, everything()) %>%
  select(-geno_risk_1, -geno_risk_2, -geno1_class, -geno2_class) %>%
  rename(`Genotype Risk` = geno_risk, `Genotype Class` = geno_class)

@@

## Table 2: Participant Demographics by Genotype Class

In the table below "MF" stands for minimal function, "RF" for residual function, and "U" indicates unknown.

```{r results='asis'}
summary(tableby(`Genotype Class` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
  data = df, control = mycontrols
))
```

## Table 3: Participant Demographics by Genotype Risk

In the table below "M" stands for mild, "S" for severe, and "U" indicates unknown.

```{r results='asis'}
summary(tableby(`Genotype Risk` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
  data = df, control = mycontrols
))
```

@@

#### By diagnosis

@@

#### By genotype class

```{r}
#| warning: false
clp <- ggplot(df, aes(x = `Genotype Class`, y = Cl)) +
  geom_boxplot() +
  theme_bw()
clp
```

#### By genotype risk

```{r}
#| warning: false
clp <- ggplot(df, aes(x = `Genotype Risk`, y = Cl)) +
  geom_boxplot() +
  theme_bw()
clp
```

@@

#### By genotype class

```{r}
#| warning: false
nap <- ggplot(df, aes(x = `Genotype Class`, y = Na)) +
  geom_boxplot() +
  theme_bw()
nap
```

#### By genotype risk

```{r}
#| warning: false
nap <- ggplot(df, aes(x = `Genotype Risk`, y = Na)) +
  geom_boxplot() +
  theme_bw()
nap
```

@@

#### By genotype class

```{r}
#| warning: false
ratiop <- ggplot(df, aes(x = `Genotype Class`, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  theme_bw() +
  ylab("Na:Cl") +
  xlab("Modulator")
ratiop
ratiop <- ggplot(df, aes(x = `Genotype Class`, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  ylim(0, 2) +
  theme_bw() +
  ylab("Na:Cl") +
  xlab("Modulator")
ratiop
```

#### By genotype risk

```{r}
#| warning: false
ratiop <- ggplot(df, aes(x = `Genotype Risk`, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  theme_bw() +
  ylab("Na:Cl") +
  xlab("Modulator")
ratiop
ratiop <- ggplot(df, aes(x = `Genotype Risk`, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  ylim(0, 2) +
  theme_bw() +
  ylab("Na:Cl") +
  xlab("Modulator")
ratiop
```
